---
title: "R programming Assignment 3"
author: "Me"
date: "12/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#1 Plot the 30 day mortality rate. 
```{r}
library(tidyverse)
outcome = read.csv('outcome-of-care-measures.csv', colClasses = 'character')
dim(outcome)
head(outcome)
names(outcome) %>%
  head()
```

To make a simple histogram of the 30 day death rates from heart attack (column 11) run
```{r}
outcome[,11] = as.numeric(outcome[,11])
hist(outcome[,11])
```

#2 Finding best hospital in the state.
best takes two argument, abbreviated name and outcome name and returns the hospital that has the best lowest 30 day mortality for specified outcome. 

```{r}
best = function(state, outcome){
  #read outcome data
  outcome_data = fread('outcome-of-care-measures.csv')
  
  #check that the outcomes and state are valid
  state_select = state
  
  if(!state %in% (outcome_data %>% distinct(State) %>% .$State)) stop('Invalid State')
  if(!outcome %in% c('heart attack','heart failure','pneumonia')) stop('Invalid Outcome')
  
  #clean the name
  outcome = janitor::make_clean_names(outcome)
  
  #clean col names
  outcome_data = janitor::clean_names(outcome_data)
  
  #get selected state
  outcome_data = outcome_data[state == state_select]
  
  #select only the needed column
  outcome_data = outcome_data %>%
    select(state, hospital_name, paste0('hospital_30_day_death_mortality_rates_from_',outcome)) %>%
    filter(complete.cases(.))
  
  #convert to numeric
  outcome_data[,3]=  as.numeric(unlist(outcome_data[,3]))
  outcome_data[order(outcome_data[,3]), 'hospital_name'][1]
}

best('NY','hert attack')
```

#Ranking hospitals by outcome.

Three arguments state, outcome, and ranking.
Would return a character vector containing name of hospital with the nth lowest death rate. can take argument as best or worst. 
If n > number of hospitals should return NA

```{r}
rankhospital = function(state, outcome, num = 'best'){
  #read outcome data
  outcome_data = fread('outcome-of-care-measures.csv')
  
  #check that the outcomes and state are valid
  state_select = state
  
  if(!state %in% (outcome_data %>% distinct(State) %>% .$State)) stop('Invalid State')
  if(!outcome %in% c('heart attack','heart failure','pneumonia')) stop('Invalid Outcome')
  
  #clean the name
  outcome = janitor::make_clean_names(outcome)
  
  #clean col names
  outcome_data = janitor::clean_names(outcome_data)
  
  #get selected state
  outcome_data = outcome_data[state == state_select]
  
  #select only the needed column
  outcome_data = outcome_data %>%
    select(state, hospital_name, paste0('hospital_30_day_death_mortality_rates_from_',outcome))
  
  #convert to double
  outcome_data[,3]=  as.numeric(unlist(outcome_data[,3]))
  
  #order based from mortality rates
  #and create ranking
  
  outcome_data = outcome_data[complete.cases(outcome_data),]
  outcome_data = outcome_data %>%
    mutate(ranking = rank(outcome_data[,3], ties.method = 'last'))
  
  #order per rank 
  outcome_data = outcome_data[order(outcome_data[,4]),]
  
  #return values
  if (num == "best"){
    return(outcome_data[1,'hospital_name'])
  }
  
  if (num == "worst"){
    return(outcome_data[nrow(outcome_data),'hospital_name'])
  }
  
  return(outcome_data[num,'hospital_name'])
  
}

rankhospital('MD','heart attack','worst')
rankhospital('MN','heart attack',5000)
rankhospital('TX','heart failure', 4)
```


# Rank all

Check the rank and get from all states

```{r}
rankall = function(outcome, num = 'best'){
  #read outcome data
  outcome_data = fread('outcome-of-care-measures.csv')
  
  #check that the outcome is valid
  
  if(!outcome %in% c('heart attack','heart failure','pneumonia')) stop('Invalid Outcome')
  
  #clean the name
  outcome = janitor::make_clean_names(outcome)
  
  #clean col names
  outcome_data = janitor::clean_names(outcome_data)
  
  
  #select only the needed column
  outcome_data = outcome_data %>%
    select(state, hospital_name, paste0('hospital_30_day_death_mortality_rates_from_',outcome))
  
  #convert to double
  outcome_data[,3]=  as.numeric(unlist(outcome_data[,3]))
  col_name = names(outcome_data[,3])
  
  outcome_data = outcome_data %>%
    filter(complete.cases(.)) %>%
    arrange(hospital_name) %>%
    group_by(state) %>%
    mutate(ranking = rank(get(col_name), ties.method = 'first')) %>%
    arrange(state, ranking, hospital_name)
  
  all_states = outcome_data %>% distinct(state)
  
  #get ranking = 1
  if(num == 'best'){
      output = outcome_data %>%
      filter(ranking == 1)
    
      output = all_states %>%
        left_join(output)
      
    return(output)
  }
  
  #get the last row
  if(num == 'worst'){
    output = outcome_data %>%
      group_by(state) %>%
      arrange(get(col_name)) %>%
      slice(n())
    
    output = all_states %>%
        left_join(output)
    
    return(output)
  }
  
  #get ranking == num
  else{
  
  output = outcome_data %>%
      filter(ranking == num)
    
      output = all_states %>%
        left_join(output)
    
    return(output)
  
  }
}


```





