---
title: "Week 3 - Looping functions and debugging"
author: "Me"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


lapply - over a list and evaluate functions
sapply - same as lapply but simplify the result
apply - apply a function over margins of array
tapply - apply a function over subsets of vector
mapply - multivariate version of lapply

split is conjuction with lapply

```{r}
lapply
```

lapply always returns a list, regardless of the input
```{r}
library(purrr)
x = list(a = 1:5, b = rnorm(10))
y = data.frame(a = 1:5, b = rnorm(10))
lapply(y, mean)
map_df(y, mean)
```

```{r}
x = list(a = 1:4, b = rnorm(10), c = rnorm(20,1),  d= rnorm(100,5))
lapply(x, mean)
```

```{r}
x = 1:4
lapply(x, runif)
```

```{r}
x = 1:4
lapply(x, runif, min = 0, max = 10)
```

Lapply and friends makes heavy use of anonymous functions
```{r}
x = list(a = matrix(1:4,2,2), b = matrix(1:6, 3,2))
lapply(x, function(elt) elt[,1])
```
Sapply will try to simplify results of lapply
- if result is a list where every element is length 1 then a vector is returned
- if result is a list where every element is a vector of same length, a matrix is returend
- if cant figure out, a list is returned


Apply
- evaluate a function over margins of array

```{r}
str(apply)
```
x - array
margin - integer vector indicating which margins shoould be retained
FUN - functions
... - other arguments

```{r}
x = matrix(rnorm(200), 20 , 10)
apply(x, 2, mean) #col
apply(x,1, mean) #row
```

For simple functions you can use 
rowSums
rowMeans,
colSums
colMeans

```{r}
colMeans(x)
colMeans(as.data.frame(x))
```
Other ways to use apply
Quantiles of the rows of matrix
```{r}
x = matrix(rnorm(200), 20, 10)
apply(x, 1, quantile, probs = c(.25, .75))
```

Average matrix in an array

this will collapse the 3rd dimension and keep 1 and 2
```{r}
a = array(rnorm(2 * 2 * 10), c(2,2,10))
apply(a, c(1,2), mean)
rowMeans(a, dims = 2)
```
Mapply - applies a function in parallel over set of arguments
```{r}
str(mapply)
```
FUN - function to apply
... functions to apply over
MoreArgs = list of other arguments to fun
Simplify = whether the results should be simplified


```{r}
list(rep(1,4), rep(2,3), rep(3,2), rep(4,1))
mapply(rep, 1:4, 4:1)
```
```{r}
noise = function(n, mean, sd){
  rnorm(n, mean, sd)
}

noise(5,1,2)
noise(1:5, 1:5, 2)
```

```{r}
list(noise(1,1,2), noise(2,2,2), noise(3,3,2), noise(4,4,2), noise(5,5,2))
```

tapply - used to apply over a subset of vector
```{r}
str(tapply)
```
X - vector
INDEX - a factor or list of factors 
FUN - function to be applied
... other arguments
simplify - should we simplify?

```{r}
x = c(rnorm(10), runif(10), rnorm(10,1))
f = gl(3,10)
tapply(x, f, mean)
```
If you dont simplify you get a 
```{r}
tapply(x, f, mean, simplify = F)
```
```{r}
str(split)
```

x - vector or list
f - factor or list of factors
drop = indicates whether empty factors should be dropped

```{r}
x = c(rnorm(10), runif(10), rnorm(10,1))
f = gl(3,10)
split(x, f)
```

```{r}
lapply(split(x,f), mean)
```

```{r}
library(datasets)
head(airquality)
```

```{r}
sapply(split(airquality$Ozone, airquality$Month), mean, na.rm = T)
```
```{r}
s = split(airquality, airquality$Month)
sapply(s, function(x) colMeans(x[,c('Ozone','Solar.R','Wind')], na.rm = T))
```

Splitting on more than 1 level
```{r}
x = rnorm(10)
f1 = gl(2,5)
f2 = gl(5,2)
f1
f2

interaction(f1,f2)
```
```{r}
str(split(x, list(f1,f2), drop = T))
```

Debugging

Basic warning
```{r}
log(-1)
```

Something is wrong
```{r}
printmessage = function(x){
  if (x > 0){
    print('x is greater than zero')
  }
  else{
    print('x is less than or equal to zero')
  }
  invisible(x)
}

printmessage(1)
printmessage(NA)
```
```{r}
printmessage2 = function(x){
  if(is.na(x)){
    print('x is a missing value')
  }
  else if (x > 0){
    print('x is greater than zero')
  }
  else{
    print('x is less than or equal to zero')
  }
}

x = log(-1)
printmessage2(x)
```

Traceback - where in the series of function the error occured
debug - give a function as argument and flags a function for debug mode
browser - allows to browser call anywhere in the code.
trace - insert debugging code in the function.
recover - error handler function. rather than getting the console back, the R interpreter will stop execution of that function.

```{r}
mean(x)
traceback()
```
Will give you most recent error.

```{r}
lm(y ~ x)
traceback()
```
```{r}
debug(lm)
lm(y~x)
```

recover
```{r}
options(error = recover) #set global option
read.csv('nosuchfile')
```


##quiz

```{r}
library(datasets)
library(data.table)
data(iris)
?iris
iris = as.data.table(iris)
class(iris)
```


```{r}
iris[Species == 'virginica', .(round(mean(Sepal.Length)))]
```

```{r}
apply(iris[,1:4], 2, mean)
```

```{r}
data("mtcars")
```

Calculate average mpg by number of cylinders
```{r}
tapply(mtcars$mpg, mtcars$cyl,mean) #correct
split(mtcars, mtcars$cyl) #incorrect
apply(mtcars, 2, mean) #incorrect
tapply(mtcars$cyl, mtcars$mpg, mean) #incorrect
sapply(split(mtcars$mpg, mtcars$cyl), mean) #correct
mean(mtcars$mpg, mtcars$cyl)#
lapply(mtcars, mean)
with(mtcars, tapply(mpg, cyl, mean))
```

```{r}
mtcars_dt = as.data.table(mtcars)
cyl_means = mtcars_dt[, .(mean_cols = mean(hp)), by = cyl]
round(abs(cyl_means[cyl == 4, mean_cols] - cyl_means[cyl == 8, mean_cols]))
```
```{r}
debug(ls)
ls()
```












