---
title: "Week2 Notes"
author: "Me"
date: "12/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

two samples of if else
```{r}
x = 3
if (x>3){
  y = 10
} else {
  y = 0
}

x = 4
y = if(x > 3) {
  10
}
else {
  0
}
```


Different ways to use for loops
```{r}
x = c('a','b','c','d')

for (i in 1:4){
  print(x[i])
}

for (i in seq_along(x)){
  print(x[i])
}

for (letter in x){
  print(letter)
}

for (i in 1:4) print(x[i])
```

Nested for loops
```{r}
x = matrix(1:6, 2, 3)
for (i in seq_len(nrow(x))){
  for(j in seq_len(ncol(x))){
    print(x[i,j])
  }
}
```

While
```{r}
count = 0
while (count < 10){
  print(count)
  count = count + 1
}
```

Repeat, next, break

```{r}
for (i in 1:100){
  if (i <= 20){
     #skip first 20 iterations
    next
  }
  #do something
}
```


```{r}
add_two = function(x,y){
  x + y
}

add_two(1,4)
```

```{r}
above_10 = function(x){
    #subset of x bigger than 10
    use = x > 10
    x[use]
}

v = c(1,10,2,11,14,5)

above_10(v)

above = function(x, n){
  #bigger than n
  use = x > n
  x[use]
  
}

above(v, 4)


```


```{r}
x = matrix(1:20, 4, 5)

mean_matrix = function(x, removeNA = T){
  nc = ncol(x)
  mean_vals = vector(mode = 'integer', length = nc)
  for(i in 1:nc){
    mean_vals[i] = mean(x[,i], na.rm = removeNA)
  }
  mean_vals
}

mean_matrix(airquality, T)
```
```{r}
args(lm)
```

Lazy evaluation
```{r}
f = function(a,b){
  a^2
}
f(2,2)
```

```{r}
f = function(a,b){
  print(a)
  print(b)
}

f(45)
```


... argument
```{r}
myplot = function(x,y , type = 'l', ...){
  plot(x, y, type = type, ...)
}
```

Will replicate original functions

```{r}
args(paste)
```
```{r}
args(cat)
```

```{r}
args(paste)

paste('a','b', sep = ':')
paste('a','b', se = ':')
```

```{r}
search()
```

This function returns another function
```{r}
make.power = function(n){
  pow = function(x){
    x^n
  }
  pow
}
```

```{r}
cube = make.power(3)
square = make.power(2)

cube(3)
square(3)
```

Exploring a functions environment
```{r}
ls(environment(cube))
get('n',environment(cube))
```
Lexical vs dynamic scoping
```{r}
y = 10

f = function(x){
  y = 2
  y^2 +g(x)
}

g = function(x){
  x*y
}

f(3)
```
Value of y in g is where the function was defined so 10

With dynamic scoping, the y is looked up from where the function was called or calling environment. 

In R, the calling environment is known as the parent frame, so the value of y is 2


WHen a function is defined in the global environment and subsequently called from the global environment, then defining the environment and calling environment are the same. This can sometimes give the appearance of dynamic scoping.

```{r}
g = function(x){
  a = 3
  x+a+y
}

g(2) #error
y = 3
g(2)
```

Maximizing a normal likelihood
Write a constructor function
```{r}
make.NegLogLik = function(data, fixed = c(F,F)){
  params = fixed
  function(p){
    params[!fixed] = p
    mu = params[1]
    sigma = params[2]
    a = -.5*length(data)*log(2*pi*sigma^2)
    b = -.5*sum((data-mu)^2)/(sigma^2)
    -(a+b)
  }
}
```


```{r}
set.seed(1)
normals = rnorm(100,1,2)
nll = make.NegLogLik(normals)

nll
ls(environment(nll))
```

```{r}
optim(c(mu = 0, sigma = 1), nll)$par
```

Fixing sigma = 2
```{r}
nll = make.NegLogLik(normals, c(F, 2))
optimize(nll, c(-1,3))$minimum
```

```{r}
x = as.Date('1970-01-01')
x
unclass(x) #days after 19700101
```

Posixlt
```{r}
x= Sys.time()
x
p = as.POSIXlt(x)
names(unclass(p))
p$sec
quarters(p)

```

posixct
```{r}
x = Sys.time()
x
unclass(x)
p = as.POSIXlt(x)
p$sec
```

```{r}
cube <- function(x, n) {
        x^3
}

cube(3)
```

```{r}
x <- 1:10
if(x > 5) {
        x <- 0
}
```

```{r}
f <- function(x) {
        g <- function(y) {
                y + z
        }
        z <- 4
        x + g(x)
}

z <- 10
f(3)
```

```{r}
x <- 5
y <- if(x < 3) {
        NA
} else {
        10
}
```




