---
title: "Week 3 Programming Assignment"
author: "Me"
date: "12/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Creates a special vector which is a list containing function to
1. set value of a vector
2. get value of vector
3. set value of mean
4. get value of mean
```{r}
make_vector = function(x = numeric()){
  m <- NULL
  set = function(y){
    x <<- y 
    m <<- NULL
  }
  get = function() x
  setmean = function(mean) m <<- mean
  getmean = function() m
  list(set = set, get = get, setmean = setmean, getmean = getmean)
}
```

This function calcualtes the mean of the special vector created with the above function. However, it first checks to see if thea mean is already calculated. If so it gets the mean from the cache and skips computation. Otherwise, it calculates the mean and set value of mean in the cache via setmean

```{r}
cachemean = function(x, ...){
  m <- x$getmean()
  if(!is.null(m)){
    message('getting cached data')
    return(m)
  }
  data = x$get()
  m = mean(data, ...)
  x$setmean(m)
  m
}
```


```{r}
x = make_vector(c(1,2,3))
x$setmean(4)
x$getmean()
x$set(c(1,2,4,5,6))
x$getmean()

cachemean(x)
x$getmean()
x$setmean(45)
x$getmean()
cachemean(x)
x$get()
x$getmean()
x$set(c(5,5))
x$get()
x$getmean()
cachemean(x)
x$getmean()
```

```{r}
makeCacheMatrix <- function(x = matrix()) {
  #initialize null inverse
  i <- NULL
  
  #set function for setting the matrix
  set <- function(y){
      
      x <<- y
      #if a matrix is set, make the inverse NULL
      i <<- NULL
  }
   
  #set function to get matrix 
  get <- function() x
  
  ##set inverse of the matrix using an input
  set_inverse <- function(inverse) i <<- inverse
  
  
  #get the inverse
  get_inverse <- function() i
    
  #return list of functions
  list(set = set, get = get, set_inverse = set_inverse, get_inverse = get_inverse)
  
}
```

```{r}
cacheSolve <- function(x, ...) {
    # get the inverse value to m
    m <- x$get_inverse()
    
    #check if m is not null then just return it
    if(!is.null(m)) {
        message("getting cached data")
        return(m)
    }
    
    #if there is no inverse get the data
    data <- x$get()
    
    #solve the inverse matrix
    m <- solve(data)
    
    #set the inverse using set_inverse
    x$set_inverse(m)
    m
    
}
```

```{r}
sample_mat = matrix(c(4,3, 3,2), 2,2)
solve(sample_mat) %*% sample_mat
sample_set = matrix(c(5,5, 5,5), 2,2)
```

```{r}
x = makeCacheMatrix(sample_mat)
x$set_inverse(sample_set)
x$get_inverse()

cacheSolve(x)
x$set(sample_mat)
x$get_inverse()
x$set(sample_mat)
cacheSolve(x)
x$get_inverse()
x$get() %*% x$get_inverse()
```

```{r}
makeVector <- function(x = numeric()) {
        m <- NULL
        set <- function(y) {
                x <<- y
                m <<- NULL
        }
        get <- function() x
        setmean <- function(mean) m <<- mean
        getmean <- function() m
        list(set = set, get = get,
             setmean = setmean,
             getmean = getmean)
}
```

```{r}
cachemean <- function(x, ...) {
        m <- x$getmean()
        if(!is.null(m)) {
                message("getting cached data")
                return(m)
        }
        data <- x$get()
        m <- mean(data, ...)
        x$setmean(m)
        m
}
```

```{r}
x = makeVector(c(1,2,4,5,6))
x$get()
x$getmean()
x$setmean()
```


```{r}
library(stringr)
character = '430bbdd9e0150b9413cfd3b397455378bf1b42f4'
str_length(character)
```

















