---
title: "Week 4 Simulation and Profiling"
author: "Me"
date: "12/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
str(str)
```
```{r}
str(lm)
```
```{r}
x = rnorm(100, 2, 4)
summary(x)
str(x)
```

```{r}
f = gl(40,10)
str(f)
```

Simulation

rnorm = random normal variates with given mean and sd
dnorm = normal probability density at a point or vector points
pnorm = evaluate cumulative distribution function for a normal distribution
rpois = generate poisson variates

d - density
r - random number generation
p - cumulative distribution
q - quantile function

```{r}
x = rnorm(10)
x = rnorm(10,20,2)
summary(x)
```

Generating poisson data
```{r}
rpois(10,1)
rpois(10,2)
ppois(2,2)
```
Simulate from a linear model
```{r}
set.seed(20)
x = rnorm(100)
e = rnorm(100, 0, 2)

y = .5 + 2*x + e

summary(y)
plot(x,y)
```


X is binary
```{r}
set.seed(10)
x = rbinom(100,1 ,.5)
e = rnorm(100,0 ,2)
y = .5 + 2 * x + e
plot(x,y)
```

Poisson distribution where Bo = .5 and B1 = .3
```{r}
set.seed(1)
x = rnorm(100)
log.mu = .5 + .3 *x
y = rpois(100, exp(log.mu))
plot(x,y)
```

Random sampling
```{r}
set.seed(1)
sample(1:10, 4)

sample(letters, 5)

sample(1:10, replace = T)
```
Elapsed time > user time
```{r}
system.time(readLines("http://www.jhsph.edu"))
```

elapsed time < user time
```{r}
hilbert = function(n){
  i <- 1:n
  1 / outer(i - 1, i , '+')
}

x = hilbert(1000)
system.time(svd(x))
```

Rprof() starts profiler in R
summaryRprof() summarizes output from RProf()

Do not use system.time or Rprof() together

SummaryProf() tabulates R profiler output and calculates how much time is spend in each function
Two methods for normalizing data
- 'by.total' divides the time n each function over total run time
- 'by.self' does same but first subtracts out time spent  in functions above the call stack

Quiz

```{r}
set.seed(1)
rpois(5, 2)
```
```{r}
set.seed(10)
x <- rep(0:1, each = 5)
e <- rnorm(10, 0, 20)
y <- 0.5 + 2 * x + e

plot(x,y)
```















